\chapter{\code{parallel.py}} % Main chapter title

\label{Chapter7} % For referencing the chapter elsewhere, use \ref{Chapter1} 

%----------------------------------------------------------------------------------------
\section{\code{Run\_Parallel\_Model}}
Function which runs the full parallel model.

It takes the following inputs:

\begin{itemize}
\item \code{Parallelisation}: String containing the choice of parallelisation.
\item \code{Scheme}: String containing the name of the numerical scheme.
\item \code{Mesh\_File}: A string containing the full location of the mesh.
\item \code{Num\_Modes}: Integer containing the value of \(N_Z\)
\item \code{P}: The degree of the polynomial.
\item \code{Num\_Constants}: The number of constants that were used in the model calibration.
\item \code{Fit}: The values for the constants in the calibration, measured in FLOPs.
\item \code{BW\_Node\_To\_Node}: Float containing the bandwidth between two nodes.
\item \code{LAT\_Node\_To\_Node}: Float containing the latency between two nodes.
\item \code{BW\_Socket\_To\_Socket}: Float containing the bandwidth between two sockets on the same node.
\item \code{LAT\_Socket\_To\_Socket}: Float containing the latency between two sockets on the same node.
\item \code{BW\_Core\_To\_Core}: Float containing the bandwidth between two sockets on the same socket.
\item \code{LAT\_Core\_To\_Core}: Float containing the latency between two sockets on the same socket.
\item \code{Num\_Core\_Per\_Socket}: Integer containing the number of cores per socket.
\item \code{Num\_Sock\_Per\_Node}: Integer containing the number of sockets per node.
\item \code{PROC\_TOT}: Integer containing the total number of processes.
\item \code{Pressure}: Dictionary containing the pressure CG iteration counts for each plane number.
\item \code{Velocity\_1}: Dictionary containing the velocity 1 CG iteration counts for each plane number.
\item \code{Velocity\_2}: Dictionary containing the velocity 2 CG iteration counts for each plane number.
\item \code{Velocity\_3}: Dictionary containing the velocity 3 CG iteration counts for each plane number.
\end{itemize}

It returns the following outputs:

\begin{itemize}
\item \code{PROC\_Z}: List of the possible \(R_Z\).
\item \code{PROC\_XY}: List of the possible \(R_{XY}\).
\item \code{Total}: List containing the time predictions for the corresponding \(R_Z\) and \(R_{XY}\).
\end{itemize}

It sends the following to the \code{stdout}:

\begin{itemize}
\item None
\end{itemize}

Additional Output:
\begin{itemize}
\item Plot showing predicted length of time step.
\end{itemize}

%----------------------------------------------------------------------------------------

\section{\code{Run\_Parallel\_Comparison}}
Function to compare the model to the data.

It takes the following inputs:

\begin{itemize}
\item \code{Loc\_Parallel\_Timing\_Files}: A string containing the location of where to find the parallel timing files to produce the plots of the model versus data.
\item \code{Parallelisation}: String containing the choice of parallelisation.
\item \code{PROC\_Z}: Integer containing the value for \(R_Z\)
\item \code{PROC\_XY}: Integer containing the value for \(R_{XY}\)
\item \code{Total}: List containing the time predictions for the corresponding \(R_Z\) and \(R_{XY}\).
\end{itemize}

It returns the following outputs:

\begin{itemize}
\item None
\end{itemize}

It sends the following to the \code{stdout}:

\begin{itemize}
\item The mean, standard deviation and differences of the difference between the model and the data.
\end{itemize}

Additional Output:
\begin{itemize}
\item Plot showing predicted length of time step vs data.
\end{itemize}

%----------------------------------------------------------------------------------------


